---
title: "Reproducible Research Project 1"
author: "Joseph"
date: "April 4, 2016"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```



###Load the data set and the plotting library:

```{r dataset}
setwd("C:/Users/JOSEPH/Documents/DataScience/RepResearch")
library(ggplot2)
ds <- read.csv("activity.csv")
head(ds)
dim(ds)
```
##Mean Total Number of Steps taken per day

We aggregate steps by each date:

```{r TotStep}
stepTot <- aggregate(steps~date,ds,sum,na.rm=TRUE)
dim(stepTot)
head(stepTot)
```
###Histogram of the total number of steps taken each day

```{r stepTotal, echo=TRUE}
ggplot(stepTot, aes(x = steps)) +geom_histogram(fill = "purple", binwidth = 1000) +
 labs(title = "Daily Total Steps", x = "Total Steps", y = "Frequency")
```

####Mean total number of steps taken per day:
```{r mean}
stepMean <-  mean(stepTot$steps, na.rm=TRUE)
stepMean
```
####Median total number of steps taken per day:
```{r median}
stepMedian <-  median(stepTot$steps, na.rm=TRUE)
stepMedian
```
##Daily Activity Pattern

Daily Steps Averages:
```{r avg}
stepAvg <- aggregate(steps~interval,ds,mean,na.rm=TRUE)
head(stepAvg)
summary(stepAvg)
```
###Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days

Let us plot Average Daily Steps
```{r plotAvg, echo=TRUE}
ggplot(stepAvg, aes(x =interval , y=steps)) +
  geom_line(size=1) +
  labs(title = "Average of Daily Steps", x = "Interval", y = "Avg. Steps per day")
```
###5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps
```{r maxinterval}
stepAvg[which.max(stepAvg$steps),]$interval
```
##Imputing missing values

Missing data will be replaced with the mean for that day
```{r newds}
dsNew <- ds
notEx <- is.na(ds$steps)
repl <- tapply(dsNew$steps, dsNew$interval, mean, na.rm=TRUE, simplify = TRUE)
dsNew$steps[notEx] <- repl[as.character(dsNew$interval[notEx])]
head(dsNew)
```
A new histogram of total steps for the comlete data set:
```{r stepTotNew, echo=TRUE}
stepAvgNew <- aggregate(steps~interval,dsNew,mean,na.rm=TRUE)
ggplot(stepTot, aes(x = steps)) +geom_histogram(fill = "blue", binwidth = 1000) +
  labs(title = "Daily Total Steps (missing data to median)", x = "Total Steps", y = "Frequency")
```
##Differences in activity patterns between weekdays and weekends

We add field wDay to the data frame to distinguish week ends from week days
```{r newFld}
dsNew$date <- as.Date(dsNew$date)
#create a vector of weekdays
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
head(dsNew)
dsNew$wDay <- factor((weekdays(dsNew$date) %in% weekdays1), 
                   levels=c(FALSE, TRUE), labels=c('weekend', 'weekday') )

##Reorder columns
dsNew<- dsNew[, c("wDay","date", "interval", "steps")]
head(dsNew)
```
Aggregate the data, split by Interval and wDay
```{r wDay}
stepAvgNewW <- aggregate(steps~interval+wDay,data=dsNew,mean,na.rm=TRUE)
head(subset(stepAvgNewW,wDay=="weekend"))
head(subset(stepAvgNewW,wDay=="weekday"))
```
Plot comparison, Week End vs. Week Day
```{r Comp, echo=TRUE}
ggplot(stepAvgNewW, aes(x =interval , y=steps, color=wDay)) +
  geom_line() +
  labs(title = "Avg. Daily Steps Week Day vs Week End", x = "Interval", y = "Num of Steps") +
  facet_wrap(~wDay, ncol = 1, nrow=2)
```
